---
title: "p8105_hw3_jt3387"
output: github_document
---
  
```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

Load the dataset:

```{r}
data("instacart")
instacart = instacart %>% 
  as_tibble(instacart)
```
  
Brief discription of the dataset:

- This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

- The dataset contains following variables: ``r names(instacart)``, and some key variables are: `order_id` represents the order identifier, `product_id` represents the product identifier, `user_id` represents the customer identifier, `aisle_id` represents the aisle identifier, `department_id` represents the department identifier.
  
- In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

- In total, there are 134 aisles, and fresh vegetables and fresh fruits aisles are the most items ordered from.

Make a plot that shows the number of items ordered in each aisle.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


