---
title: "p8105_hw3_jt3387"
output: github_document
---
  
```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

Load the dataset:

```{r}
data("instacart")
instacart = instacart %>% 
  as_tibble(instacart)
```
  
Brief discription of the dataset:

- This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

- The dataset contains following variables: ``r names(instacart)``, and some key variables are: `order_id` represents the order identifier, `product_id` represents the product identifier, `user_id` represents the customer identifier, `aisle_id` represents the aisle identifier, `department_id` represents the department identifier.
  
- In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

- In total, there are 134 aisles, and fresh vegetables and fresh fruits aisles are the most items ordered from.

Make a plot that shows the number of items ordered in each aisle.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. 

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r message = FALSE}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

Load, tidy, and wrangle the data.

```{r message = FALSE}
accel_df <- 
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_prefix = "activity_",
    names_to = "minute",
    values_to = "activity"
  ) %>% 
  mutate(
    minute = as.numeric(minute),
    dow = ifelse(day %in% c('Saturday', 'Sunday'), "weekend", "weekday")
  ) %>% 
  relocate(week, day_id, day, dow)
```

Brief discription of the resulting dataset:

- The data contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns.

- There are 7 variables in the data set: `week` represents the week number from 1 to 5, `day_id` shows the id of each day ranging from 1 to 35, `day` shows what day of the week it is, `dow` shows whether the day is a weekday or weekend, `minite` represents every minute within a 24-hour day ranging from 1 to 1440, `activity` shows the activity counts for each minute of a 24-hour day starting at midnight, and its min value is ``r min(pull(accel_df, activity))`` and max value is ``r max(pull(accel_df, activity))``.

Analyses of the data focus on the total activity over the day.

```{r message = FALSE}
total_act <- 
  accel_df %>% 
  group_by(week, day) %>% 
  summarize(sum_activity = sum(activity))
```

Create a table showing the totals.

```{r}
total_act %>% 
  pivot_wider(
    names_from = "day",
    values_from = "sum_activity"
  ) %>% 
  relocate(1, 5, 3, 7, 8, 6, 2, 4) %>% 
  knitr::kable(
    caption = "Total activity over a day",
    align = "l"
  )
```